---
title: "Portfolio4"
author: "Liv Tollånes"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Heart rate, respiration and interpersonal coordination

Physiological data (here heart rate [variability], and respiration) are increasingly popular. Historically treated as pernicious noise to be regressed out of neuro-imaging data, there is now increasing research on how these signals tell us something important about cognition and beyond being just a signal of cognitive processes also impact them in interesting ways. Advanced sport science, and the quantified self movement (closely followed by marketing and communication) have hailed continuous physiological tracking as a powerful way to access and modify attitudes, habits, and performance. Further, as team coordination (in the military, in decision processes and organizational contexts) is more and more in focus, research has attempted to measure how interpersonal coordination between physiological systems might tell us something important about e.g. emotional and cognitive coordination. See references in the reading list for more on this.

In this assignment, you will learn to:
- pre-process physiological data (and grow further your mad R skills)
- model the continuous interdependence between two signals (using a multilevel model as proxy for a dynamical system approach)
- conservatively assess the presence of coordination between to signals in a controlled context

This assignment has two parts. The first part familiarizes you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. N.B. The data are collected by students from previous years (Study 1 - 4). Note that synchronous and turn-taking are the same across all four studies, but the third condition is different: in the first year it was self-paced joint reading; in the second to fourth years it was the tv-series conversation.

## Let's get started

### Exploring physiological signals
The data files can be found here: https://www.dropbox.com/sh/bvvk7t3fvsplh9o/AADM6q4WrtXKvSwH5aAO1umta?dl=0

- Choose one pair (one pair, three conditions, three files)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal.
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3). There are also smarter packages, like cowplot and ggpubr.
- Can you eye-ball which condition if any displays more physiological coordination?

### First we read one data file and identify the procedure
- Load the file
- correctly identify all columns
- plot the data
- deal with the artifacts
- downsample the data
- Add a column for study, group, trial and condition

```{r}
# Load the libraries
pacman::p_load(dplyr, tidyverse, caret, ggplot2, groupdata2, gridExtra, cowplot, ggpubr, RHRV, stats, quantmod, Hmisc, lme4)

# Load the file
d1 <- read.csv("data/Study1_G1_T1_Synchronous.csv")


# Plotting heart rate and respiration for one pair of participants before doing anything with the data
heartrate_raw <-ggplot(data = d1) + geom_path(aes(time, HR1, color = "P1")) + geom_path(aes(time, HR2, color = "P2")) + labs(x = "time", y = "HR") + theme_classic() + ggtitle("Heart rate raw data")
respiration_raw <- ggplot(data = d1) + geom_path(aes(time, Resp1, color = "P1")) + geom_path(aes(time, Resp2, color = "P2")) + labs(x = "time", y = "Respiration") + theme_classic() + ggtitle("Respiration raw data")

heartrate_raw
respiration_raw

# It is pretty obvious from the respiration plot that the data has not been scaled yet
```


```{r}
## Remove outliers by replacing values above or below our set thresholds with mean values plus the threshold times the standard deviation of our time series
removeOuts <- function(ts, threshold){
  higher_threshold_condition <- ts > (mean(ts, na.rm = T) + (threshold*sd(ts, na.rm = T)))
  lower_threshold_condition <- ts < (mean(ts, na.rm = T) - (threshold*sd(ts, na.rm = T)))
  ts[higher_threshold_condition] <- mean(ts, na.rm = T) + (threshold *sd(ts, na.rm = T))
  ts[lower_threshold_condition] <- mean(ts, na.rm = T) - (threshold *sd(ts, na.rm = T))
  return(ts)
}

threshold = 2.5

#Applying the outlier-removing function to our data
no_outliersdf <- d1 %>% 
  mutate(HR1 = removeOuts(HR1, threshold),
         HR2 = removeOuts(HR2, threshold),
         Resp1 = removeOuts(Resp1, threshold),
         Resp2 = removeOuts(Resp2, threshold))

#Plot with artifacts removed
heartrate_artrem <-ggplot(data = no_outliersdf) + geom_path(aes(time, HR1, color = "P1")) + geom_path(aes(time, HR2, color = "P2")) + labs(x = "time", y = "HR") + theme_classic() + ggtitle("Heartrate no artifacts")
respiration_artrem <- ggplot(data = no_outliersdf) + geom_path(aes(time, Resp1, color = "P1")) + geom_path(aes(time, Resp2, color = "P2")) + labs(x = "time", y = "Respiration") + theme_classic() + ggtitle("Respiration no artifacts")

#Plotting the raw data against those with artifacts removed
grid.arrange(heartrate_raw, heartrate_artrem)
grid.arrange(respiration_raw, respiration_artrem)
```



```{r}
## Scaling the variables corrected for outliers
scaled_noout <- no_outliersdf%>% 
  mutate(HR1 = scale(HR1),
         HR2 = scale(HR2),
         Resp1 = scale(Resp1),
         Resp2 = scale(Resp2))
         
### Tip: if scale() gives some issues, try the one below
#z_scale <- function(column){
#  column_c <- (column - mean(column)) / sd(column)
#}

# Plot again to check how scaled data look like
ggplot(data = scaled_noout) + geom_path(aes(time, HR1, color = "P1")) + geom_path(aes(time, HR2, color = "P2")) + labs(x = "time", y = "HR") + theme_classic() + ggtitle("Scaled HeartRate")
ggplot(data = scaled_noout) + geom_path(aes(time, Resp1, color = "P1")) + geom_path(aes(time, Resp2, color = "P2")) + labs(x = "time", y = "Respiration") + theme_classic() + ggtitle("Scaled Respiration")

```

```{r}
## Downsample
### You can use the code below. It is almost the same as the one in the video lecture.

d2 <- scaled_noout %>% group(n = 100, method = 'greedy') %>% 
  summarise(time = mean(time, na.rm = T),
            HR1 = mean(HR1, na.rm = T),
            HR2 = mean(HR2, na.rm = T),
            Resp1 = mean(Resp1, na.rm = T),
            Resp2 = mean(Resp2, na.rm = T)
            )
    
## Plot the downsampled data. You can use the code from the slides
ggplot(data = d2) + geom_path(aes(time, HR1, color = "P1")) + geom_path(aes(time, HR2, color = "P2")) + labs(x = "time", y = "HR") + theme_classic() + ggtitle("Heart rate downsampled, scaled and no outliers")
ggplot(data = d2) + geom_path(aes(time, Resp1, color = "P1")) + geom_path(aes(time, Resp2, color = "P2")) + labs(x = "time", y = "Respiration") + theme_classic() + ggtitle("Respiration downsampled, scaled, and no outliers")
```


```{r}
## Now add the group, trial, condition to the cleaned up, scaled, downsampled data
## Tip the info is in the file name
#parse filename; study, group, trial

read_filename <- function(filename) {
  filenamepath = paste("data/", filename, sep="")
  data = read_csv(filenamepath)
    labels = str_match(filenamepath, "data/Study([0-4])_G([0-9]+)_T([0-5]+)_([a-zA-Z]+).csv")
    
    #filenamesplit <- str_detect(filename, "Synchronous.csv")
    #Condition = filenamesplit
    Study = labels[2]
    Group = labels[3]
    Trial = labels[4]
    Condition = labels[5]

    
        output = data.frame(Study, Group, Trial, Condition)
    return(output)
}

d1_new <- read_filename("Study1_G1_T1_Synchronous.csv")
merged_data <- merge(d1, d1_new)

```


## Now we are ready to go to load and pre-process all files

Go through all the files (with a function passed onto map_df), check which files should be excluded, if any, and save the pre-processed time-series. This procedure is similar to what you have done in portfolio 3. You may use the code you wrote for that assignment and adjust it to this one.

A couple of tips:
- looping will be too slow for these files (remember you have ~200 000 rows in each file!). Making a function and using Map/Map_df is your salvation.
- you may want your first step after loading a file to be downsampling, so that you don't work with enormous amount of data
- each study restarts the group numbering, so you should make sure to change that (e.g. 100 * Study + Group)
- you need to make sure all the data are meaningful or something has to be removed. Plotting is your friend. E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs


```{r}
# Define a function running the loading, artifact removal, scaling, downsampling, info adding.
### function

prep <- function(filename, threshold = 2.5) {
  filenamepath = paste("data/", filename, sep="")
  l <- read_csv(filenamepath) %>% 
    rownames_to_column()
 
   #info adding
    labels = str_match(filenamepath, "data/Study([0-4])_G([0-9]+)_T([0-5]+)_([a-zA-Z]+).csv")
    Study = labels[2]
    Studynum = as.numeric(labels[2])
    Groupfactor = labels[3]
    Group = as.numeric(Groupfactor) + 100*Studynum
    Trial = labels[4]
    Condition = labels[5]
    output <- data.frame(Study, Group, Trial, Condition)
    
     if (Study == 4) {
      l <- rename(l, time = TimeMs)
     }
  
#downsampling
  l <- l %>% group(n = 100, method = 'greedy') %>%
  summarise(time = mean(time, na.rm = T),
            HR1 = mean(HR1, na.rm = T),
            HR2 = mean(HR2, na.rm = T),
            Resp1 = mean(Resp1, na.rm = T),
            Resp2 = mean(Resp2, na.rm = T)
            )

#Applying the outlier-removing function to our data
  l <- l %>% mutate(HR1 = removeOuts(HR1, threshold),
         HR2 = removeOuts(HR2, threshold),
         Resp1 = removeOuts(Resp1, threshold),
         Resp2 = removeOuts(Resp2, threshold))
  
#scaling
  l <- l %>% mutate(HR1 = scale(HR1),
         HR2 = scale(HR2),
         Resp1 = scale(Resp1),
         Resp2 = scale(Resp2))
  
    final <- merge(l, output)
    
  return(final)
}

hr = list.files(path = "data/", pattern = "csv") %>% ## NB replace with your path to the files 
    purrr::map_df(prep)

hr$.groups <-  NULL

#nrow(distinct(hr))
hr <- hr[complete.cases(hr), ]

```

```{r}
# Now we need to make sure all the data are meaningful or something has to be removed
# E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs

# plots plots plots

plots_folder <- "HR_plots/"

# Create graphing function
HR_plots <- function(df, na.rm = TRUE, ...){
  
  # create list of ID in data to loop over 
  group_list <- unique(df$Group)
  
  # create for loop to subset group ID
  for (i in seq_along(group_list)) { 
    
    subset_ID <- subset(df, df$Group==group_list[i])
    
    # create list of trials in data to loop over
    trial_list = unique(subset_ID$Trial)
    
    # create for loop to produce plot for each trial
    for (j in seq_along(trial_list)) {
      
      # create plot for each trial in subset_ID 
      plot <- 
      ggplot(subset(subset_ID, subset_ID$Trial==trial_list[j])) +
      
      geom_path(aes(time, HR1, color = "P1")) + 
      
      geom_path(aes(time, HR2, color = "P2")) +
      
      labs(x = "Time (in minutes)", y = "Heart Rate") + theme_bw()
      
      # save plots as .png
      ggsave(plot, file=paste(plots_folder,
                            "Group",group_list[i],"_trial",trial_list[j], ".png", sep=''), scale=2)
      # print plots to screen
      print(plot)
    }
  }
}

#HR_plots(hr)


plots_folder2 <- "Resp_plots/"

# Create graphing function
Resp_plots <- function(df, na.rm = TRUE, ...){
  
  # create list of ID in data to loop over 
  ID_list <- unique(df$Group)
  
  # create for loop to subset group ID
  for (i in seq_along(ID_list)) { 
    
    subset_ID <- subset(df, df$Group==ID_list[i])
    
    # create list of trials in data to loop over
    trial_list = unique(subset_ID$Trial)
    
    # create for loop to produce plot for each trial
    for (j in seq_along(trial_list)) {
      
      # create plot for each trial in subset_ID 
      plot <- 
      ggplot(subset(subset_ID, subset_ID$Trial==trial_list[j])) +
      
      geom_path(aes(time, Resp1, color = "P1")) + 
      
      geom_path(aes(time, Resp2, color = "P2")) +
      
      labs(x = "Time (in minutes)", y = "Respiration signal") + theme_bw()
      
      # save plots as .png
      ggsave(plot, file=paste(plots_folder2,
                            "Group",ID_list[i],"_trial",trial_list[j], ".png", sep=''), scale=2)
      # print plots to screen
      print(plot)
    }
  }
}

#Resp_plots(hr)



# Eyeballing the data and removing the files that does not look nice. This is to avoid removing way too many observations. One could also remove all observations where several rows after each other are identical by looking at lag and lead, but we found that this removed way too many observations for the analysis to make sense. 
# Bad observations in HR plots:

#Group205_trial1
#Group205_trial2
#Group205_trial3

#Group206_trial1
#Group206_trial2
#Group206_trial3

#Group208_trial1.
#Group208_trial2
#Group208_trial3

#Group209_trial1

#Group301_trial3

#Group402_trial1

#Group403_trial1
#Group403_trial3
#Group403_trial4
#Group403_trial5

#Group405_trial1
#Group405_trial3

#Group406_trial1
#Group406_trial2
#Group406_trial3

#Group407_trial1
#Group407_trial2
#Group407_trial5

#Group408_trial4

# Bad observation in respiration plots:

#Group101_trial1.
#Group101_trial2
#Group102_trial1
#Group102_trial2
#Group401_trial4
#Group401_trial5
#Group309_trial2




# Remove bad data in heart rate
hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==205, "NA", hr$HR1)) #Group205_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==205, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==205, "NA", hr$HR1)) #Group205_trial2
hr$HR2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==205, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==205, "NA", hr$HR1)) #Group205_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==205, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==206, "NA", hr$HR1)) #Group206_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==206, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==206, "NA", hr$HR1)) #Group206_trial2
hr$HR2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==206, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==206, "NA", hr$HR1)) #Group206_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==206, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==208, "NA", hr$HR1)) #Group208_trial1 
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==208, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==208, "NA", hr$HR1)) #Group208_trial2
hr$HR2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==208, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==208, "NA", hr$HR1)) #Group208_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==208, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==209, "NA", hr$HR1)) #Group209_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==209, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==301, "NA", hr$HR1)) #Group301_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==301, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==402, "NA", hr$HR1)) #Group402_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==402, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==403, "NA", hr$HR1)) #Group403_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==403, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==403, "NA", hr$HR1)) #Group403_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==403, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==4 & hr$Group==403, "NA", hr$HR1)) #Group403_trial4
hr$HR2 <-as.numeric( ifelse(hr$Trial==4 & hr$Group==403, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==403, "NA", hr$HR1)) #Group403_trial5
hr$HR2 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==403, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==405, "NA", hr$HR1)) #Group405_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==405, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==405, "NA", hr$HR1)) #Group405_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==405, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==406, "NA", hr$HR1)) #Group406_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==406, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==406, "NA", hr$HR1)) #Group406_trial2
hr$HR2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==406, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==406, "NA", hr$HR1)) #Group406_trial3
hr$HR2 <- as.numeric(ifelse(hr$Trial==3 & hr$Group==406, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==407, "NA", hr$HR1)) #Group407_trial1
hr$HR2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==407, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==407, "NA", hr$HR1)) #Group407_trial2
hr$HR2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==407, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==407, "NA", hr$HR1)) #Group407_trial5
hr$HR2 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==407, "NA", hr$HR2))

hr$HR1 <- as.numeric(ifelse(hr$Trial==4 & hr$Group==408, "NA", hr$HR1)) #Group408_trial4
hr$HR2 <- as.numeric(ifelse(hr$Trial==4 & hr$Group==408, "NA", hr$HR2))


# Remove bad data in respiration
hr$Resp1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==101, "NA", hr$Resp1)) #Group101_trial1
hr$Resp2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==101, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==101, "NA", hr$Resp1)) #Group101_trial2
hr$Resp2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==101, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==102, "NA", hr$Resp1)) #Group102_trial1
hr$Resp2 <- as.numeric(ifelse(hr$Trial==1 & hr$Group==102, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==102, "NA", hr$Resp1) )#Group102_trial2
hr$Resp2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==102, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==4 & hr$Group==401, "NA", hr$Resp1)) #Group401_trial4
hr$Resp2 <- as.numeric(ifelse(hr$Trial==4 & hr$Group==401, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==401, "NA", hr$Resp1)) #Group401_trial5
hr$Resp2 <- as.numeric(ifelse(hr$Trial==5 & hr$Group==401, "NA", hr$Resp2))

hr$Resp1 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==309, "NA", hr$Resp1)) #Group309_trial2
hr$Resp2 <- as.numeric(ifelse(hr$Trial==2 & hr$Group==309, "NA", hr$Resp2))


#Alternative code that looks at lead and lag in order to remove rows. A lot more observations are removed this way. 
# Remove bad data by removing values where there are several data points in a row with the exact same values -> thereby removing clipping and unnatural data
#removeArtefacts1 <- function(ts){
#  ts = ts[with(ts, c(HR1[-1]!= HR1[-nrow(ts)], TRUE)),]
#  ts = ts[with(ts, c(HR2[-1]!= HR2[-nrow(ts)], TRUE)),]
#  return(ts)
#  }
#newHR1 <- removeArtefacts1(hr)

#removeArtefacts2 <- function(ts){
#  ts = ts[with(ts, c(Resp1[-1]!= Resp1[-nrow(ts)], TRUE)),]
#  ts = ts[with(ts, c(Resp2[-1]!= Resp2[-nrow(ts)], TRUE)),]
#  return(ts)
#  }

#newdfclean <- removeArtefacts2(newHR1)

```



## Now we need to run some analysis

Let's start with a multilevel model that accounts for 
- stability (how each signal is autocorrelated)
- interpersonal dependence (each signal is dependent from the previous state of the other signal)

The data needs to be further prepared, so we can analyze both participants in the same model.
We need to turn the data into a long format:
- a column indicating own hr and one own respiration
- a column indicating other hr and one other respiration
- a column indicating change in hr from previous round and one in respiration

We can then run an analysis where change is a function of one's previous state (stability, see slides), and the other's previous state (coupling). Make sure to:
- set up the most interesting contrasts: how do these parameters vary by condition? which condition should be baseline?
- set up the right random effects.
- N.B. the model will be slow. Make sure it works on a subset of the data first!

Bonus question: what if we include an additional layer? Is my heart rate just adjusting to yours, or also to how much you are adjusting to mine?
- to start answering this we can add a column indicating the previous change in hr in the other and one in respiration
- we can then build on the previous models by also adding the previous change in the other


```{r}
# Generate a column for each: previous HR1, HR2, Resp1, Resp2. Tip: use the function Lag()
# Generate a column for each: change in HR1, HR2, Resp1, Resp2
# Make the data long, so we can analyze both participants at the same time. Tip: you can use the function gather () 
## N.B. This is a bit tricky and you might have to do it in several steps

# Making data long
df_long <- hr %>%
    pivot_longer(
      c(HR1, HR2, Resp1, Resp2),
      names_to = c(".value", "Participant"),
      names_pattern = "(Resp|HR)(\\d)"
    )

# Making "other" and "self" columns
df_long <- df_long %>% 
  group_by(Group, Trial, time) %>% 
  mutate(HR_other = ifelse(Participant == 1, Lag(HR, -1), Lag(HR, 1)),
         Resp_other = ifelse(Participant == 1, Lag(Resp, -1), Lag(Resp, 1))) %>% 
  rename(
    HR_self = HR,
    Resp_self = Resp
  )

# Computing "lead", "change", and "previous" columns
df <- df_long %>% 
  group_by(Participant) %>% 
  mutate(
  HR_self_lead = Lag(HR_self, -1),
  HR_self_change = HR_self_lead - HR_self,
  HR_other_lead = Lag(HR_other, -1),
  HR_other_change = HR_other_lead - HR_other,
  Resp_self_lead = Lag(Resp_self, -1),
  Resp_self_change = Resp_self_lead - Resp_self,
  Resp_other_lead = Lag(Resp_other, -1),
  Resp_other_change = Resp_other_lead - Resp_other,
  HR_self_previous = Lag(HR_self, 1),
  HR_other_previous = Lag(HR_other, 1),
  Resp_self_previous = Lag(Resp_self, 1),
  Resp_other_previous = Lag(Resp_other, 1)
  ) %>% subset(!is.na(time))

# Making unique participant ID
df$participantID <- as.integer(factor(with(df, paste(Group, Participant))))

#dataframe <- write.csv(df)
install.packages("readr")
library(tidyverse)

fdf <- read.csv("finaldata.csv")

# Set the most interesting contrast e.g. by defining synchronous or conversation as the baseline
# The baseline is set to be Selfpaced

library(lme4)

fdf$Condition <- factor(fdf$Condition, levels = c("SelfPaced", "TurnTaking","Conversation", "MovementGuided", "MovementCoop", "Synchronous"))


#IS THE HEART RATE OF SELF AND OTHER A GOOD PREDICTOR OF CHANGE? overall question

# Model change as a function of own and other previous state 

hr_change <- lme4::lmer(HR_self_change ~ Lag(HR_other, 1) + Lag(HR_self, 1) + (1|participantID), data = df, REML = FALSE)


# Bonus points: Add to the previous model also change in the other to see whether my adaptation is influenced by the other's adaptation.
```

```{r}
## Making models testing effects of parameters for each condition separately
model1 <- lm(HR_self_change ~ 1 + Condition + (HR_self + (HR_other-HR_self)):Condition, data = fdf)
summary(model1)
plot(model1)

# Adding random effects
model2 <- lme4::lmer(HR_self_change ~ 1 + Condition + (HR_self + (HR_other-HR_self)):Condition + (1 + HR_self | participantID) + (1 + HR_self | Group), data = fdf)
summary(model2)
plot(model2)


#Building models to predict change in heart rate and respiration before doing control tests
#A simple model for HR
HR_basic_1 <- lmerTest::lmer(HR_self_change ~ 1 + (HR_self+(HR_other-HR_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), subset(fdf, Study == "4"), REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(HR_basic_1)

#A more complex model for HR
sub_basic <- subset(fdf, Group == 101:102)

HR_basic_2 <- lmerTest::lmer(HR_self_change ~ 1 + (HR_self+(HR_other-HR_self)):Condition + (1 + (HR_self+HR_other):Condition|participantID) + (1 + (HR_self+HR_other):Condition|Group), data = sub_basic, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(HR_basic_2)

#A complex model for respiration 
RESP_basic_1 <- lmerTest::lmer(Resp_self_change ~ 1 + (Resp_self+(Resp_other-Resp_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), subset(fdf, Study == "4"), REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(RESP_basic_1)

#A more simple model for respiration 
RESP_basic_2 <- lmerTest::lmer(Resp_self_change ~ 1 + (Resp_self + (Resp_other - Resp_self)):Condition + (1 + Condition|participantID), subset(fdf, Study == "4"), REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(RESP_basic_2)


# Model comparison of heart rate models 
MuMIn::r.squaredGLMM(HR_basic_1)
MuMIn::r.squaredGLMM(HR_basic_2)
AIC(HR_basic_1)
AIC(HR_basic_2)


#Model comparison of respiartion models
MuMIn::r.squaredGLMM(RESP_basic_1)
MuMIn::r.squaredGLMM(RESP_basic_2)
AIC(RESP_basic_1)
AIC(RESP_basic_2)


# Testing the chosen models on the entire data
HR_chosen <- lmerTest::lmer(HR_self_change ~ 1 + (HR_self+(HR_other-HR_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), data = fdf, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(HR_chosen)

RESP_chosen <- lmerTest::lmer(Resp_self_change ~ 1 + (Resp_self+(Resp_other-Resp_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), data = fdf, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(RESP_chosen)
```

## Now we need to create control baselines.

First shuffled controls, then surrogate pairs.

### Creating controls: shuffled controls

Shuffled controls break the temporal dependencies of time-series by shuffling the value within one time-series. This ensures the "coordination" observed is not due to the actual values in the series and not their sequence.
Tip: sample() is your friend, but make sure to shuffle things within participant/condition and not throughout the whole dataset
 
```{r}
# Create a shuffled dataset
shuffled_data <- fdf %>% 
  group_by(participantID, Participant) %>% 
  mutate(HR_self = sample(HR_self), HR_other = sample(HR_other)) %>% 
  mutate(HR_self_change = lead(HR_self)-HR_self, HR_other_change = lead(HR_other)-HR_other, type = 0)

shuffled_data <- as.data.frame(shuffled_data)

# Concatenate it to the original dataset 
shuffled_fdf <- dplyr::bind_rows(shuffled_data, fdf)
shuffled_fdf$type <- as.factor(shuffled_fdf$type)

#MODEL FOR HEART RATE SHUFFLED

HR_chosen_shuffled <- lmerTest::lmer(HR_self_change ~ 1 + (HR_self+(HR_other-HR_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), data = shuffled_fdf, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(HR_chosen_shuffled)


#MODEL FOR RESPIRATION SHUFFLED
RESP_chosen_shuffled <- lmerTest::lmer(Resp_self_change ~ 1 + (Resp_self+(Resp_other-Resp_self)):Condition + (1 + Condition|participantID) + (1 + Condition|Group), data = shuffled_fdf, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(RESP_chosen_shuffled)


```


 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair)

```{r}
# Identify unique pairs within a given study (to keep things manageable) and create list of possible surrogate pairs (e.g. individual 1 from pair 1 and individual 2 from pair 2)

# Starting from the wide format, create "surrogate" dataset with the data from surrogate pairs
fdf$Group <- as.numeric(as.character(fdf$Group))
Groups <- unique(fdf$Group[fdf$Study == 1])
SurrogateList1 <- expand.grid (a = Groups, b = Groups)
SurrogateList1 <- subset(SurrogateList1, a != b )
Groups <- unique(fdf$Group[fdf$Study == 2])
SurrogateList2 <- expand.grid (a = Groups, b = Groups)
SurrogateList2 <- subset(SurrogateList2, a != b )
Groups <- unique(fdf$Group[fdf$Study == 3])
SurrogateList3 <- expand.grid (a = Groups, b = Groups)
SurrogateList3 <- subset(SurrogateList3, a != b )
SurrogateList <- rbind(SurrogateList1, SurrogateList2, SurrogateList3)

z1 <- subset(fdf, Group == SurrogateList$a[1])
z2 <- subset(fdf, Group == SurrogateList$b[1])
z2 <- z2[-(12352:12429),]



for (i in 1:nrow(SurrogateList)){
  x <- subset(fdf, Group == SurrogateList$a[i])
  y <- subset(fdf, Group == SurrogateList$b[i])
  group <- 800 + i
  
  for (co in c("Synchronous", "TurnTaking", "SelfPaced", "MovementGuided", "MovementCoop", "Conversation")){
    if (co %in% unique(x$Condition) & co %in% unique(y$Condition)){
      z1 <- subset(x, Condition == co)
      z2 <- subset(y, Condition == co)
    }
    if (nrow(z1) > nrow(z2)){
      z1 <- z1[-((nrow(z2)+1):nrow(z1)),]
    }
    if (nrow(z2) > nrow(z1)){
      z2 <- z2[-((nrow(z1)+1):nrow(z2)),]
    }
    w1 <- z1 %>% mutate(
      HR_other = z2$HR_other,
      Resp_other = z2$Resp_other,
      HR2_lead = z2$HR2_lead,
      Resp2_lead = z2$Resp2_lead,
      HR2_change = z2$HR2_change,
      Resp2_change = z2$Resp2_change
      )
    w1$Group <- group
    w1$Type <- "Surrogate"
    w <- w1
    if(exists("d_surrogate")){d_surrogate <- rbind(d_surrogate, w)}else{d_surrogate <- w}
  }
}

fdf$Type = "Real"
newnewdf <- rbind(fdf, d_surrogate)



#Make it into long format

Resp_model6 <- lmerTest::lmer(R_self_change ~ 0 + (R_self_lag + R_other_lag):Condition:Type + (0+(R_self_lag + R_other_lag):Condition|name), data = dd_long_sub, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

# Create models as in chunks above, but adding an interaction with the Real vs. Surrogate variable (exclude shuffled ones for simplicity)


HR_surrogate <- lmerTest::lmer(HR_self_change ~ (HR_self + (HR_other-HR_self)):Condition:Type + (1 + Condition | participantID), data = newnewdf, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))
summary(HR_surrogate)

comparison_sub <- subset(fdf, Study == "3")

RESP_chosen_surrogate <- lmerTest::lmer(Resp_self_change ~ 1 + (Resp_self+(Resp_other-Resp_self)):Condition:Type + (1 + Condition|participantID) + (1 + Condition|Group), data = comparison_sub, REML = F, control = lmerControl(optimizer = "nloptwrap", calc.derivs = F))

summary(RESP_chosen_surrogate)

```
 

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 
 - Optional: run the models and report them
